---
title: "A visualization of random effects model and fixed effects model"
author: "Miao Cai"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Random intercept model

Statistical model
---
In this section, we will set up the statistical model for a random-intercept model:
$$
\begin{aligned}
y_{ij} & = \beta_{0j} + \beta_{1}*x_{1i} +  \beta_{2}*x_{2i} + \beta_{3}*x_{3i} + \epsilon_{i}\\
\beta_{0j} & \sim N(\mu_0, \sigma_0^2)
\end{aligned}
$$

Where $i = 1, 2, \cdots, N$ is the index for individual observations, so the total sample size is $N$ in this case. $j = 1, 2, \cdots, J$ is the index for groups, so there are $J$ groups/clusters in this case. $\beta_{01}, \beta_{02}, \cdots, \beta_{0J}$ are the random intercepts in this model. $\beta_1, \beta_2, \cdots, \beta_K$ are the fixed-effects parameters, and in this case, $K = 3$. The number of predictor variables can vary, and this data generating process can be generalized using a more succinct matrix form style as:
$$
\begin{aligned}
y_{ij} & = \beta_{0j} + \mathbf{X\beta} + \epsilon_{i}\\
\beta_{0j} & \sim N(\mu_0, \sigma_0^2)
\end{aligned}
$$
where $\mathbf{X}$ is a $N\times K$ data matrix, and $\mathbf{\beta}$ is a $K \times 1$ column vector.

We assume that the random intercepts $\beta_{0j}$ has a normal distribution with two hyperparameters: mean $\mu_0$ and standard deviation $\sigma_0$.

Simulation parameters
---

### Observations and groups

- The number of groups/clusters is $J = 10$,
- The number of observations per group is generated from a homogeneous Poisson process with the mean parameter of $\lambda = 40$.

### Model parameters

- Hyperparameters: $\mu = 1, \sigma = 1$,
- Number of fixed parameters: $K = 3$,
- Fixed-effects parameter vector: $\mathbf{\beta} = [1, -1, 0.5]^\prime$
- Error term: $\epsilon_i \sim \text{Normal}(0, 5)$
- Data matrix:
  + $x_1 \sim \text{Normal}(2, 3)$,
  + $x_2 \sim \text{Poisson}(3)$,
  + $x_3 \sim \text{Gamma}(5, 5)$


Data simulation
---

```{r SimulateData, eval=TRUE}
pacman::p_load(lme4, broom, dplyr)
set.seed(666)

# Obs and groups
J = 10 # of the groups
lambda = 40 # mean number of obs per group
N_j = rpois(J, lambda) # number of obs per group
N = sum(N_j) # total number of obs

# Model parameters
mu0 = 1; sigma0 = 1 # Hyperparameters
b0_j = rnorm(J, mu0, sigma0) # Random intercepts for j-th group
b0_i = rep(b0_j, N_j) # Random intercepts for i-th obs

K = 3
b_vector = c(1, -1, 0.5)
epsilon = rnorm(N, 0, 5)

X = as.matrix(data.frame(
  x1 = rnorm(N, 2, 3),
  x2 = rpois(N, 3),
  x3 = rgamma(N, 5, 5)
))
id = rep(1:J, N_j)

# y = b0 + b1 + e
y = b0_i + X%*%b_vector + epsilon
dt = data.frame(id = id, y = y, X) # simulated data

rfit = function(fit){
 r = fit %>% 
  broom::tidy() %>% 
  mutate_if(is.numeric, round, 3)
 return(r)
}
```

Parameter estimation
===

Linear model
---

```{r}
dt %>% 
  lm(y ~ x1 + x2 + x3, data = .) %>%  
  rfit() 
```

Linear model with group-level fixed effects
---

```{r}
dt %>% 
  lm(y ~ factor(id) + x1 + x2 + x3, data = .) %>% 
  rfit() %>% 
  filter(grepl("Intercept|x", term))
```

Random-effects model
---

```{r}
dt %>% 
  lmer(y ~ x1 + x2 + x3 + (1|id), data = .) %>% 
  rfit()
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
require(ggplot2)

nDriver = 50
Hour = 10

x = seq(0, Hour, 0.1)
driverID = 1:nDriver
b0 = rnorm(nDriver, -2, 1)
b1 = rnorm(nDriver, 0.25, 0.2)

dat = data.frame(driverID = rep(driverID, each = length(x)), 
                 b0 = rep(b0, each = length(x)), 
                 b1 = rep(b1, each = length(x)), 
                 x = rep(x, nDriver))


dat$p = 1 - exp(-exp(dat$b0 + dat$b1*dat$x))

ggplot(data = dat, aes(x = x, y = p, group = driverID)) + 
  geom_line(size = 0.1) + 
  xlab("Cumulative hours of driving") + 
  ylab("Probability of at least 1 critical event") + 
  ggtitle("Random intercept and random slope model of driver fatigue")
```





