---
title: "A visualization of random effects model and fixed effects model"
author: "Miao Cai"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Random intercept model

Statistical model
---
In this section, we will set up the statistical model for a random-intercept model:
$$
\begin{aligned}
y_{ij} & = \beta_{0j} + \beta_{1}*x_{1i} +  \beta_{2}*x_{2i} + \beta_{3}*x_{3i} + \epsilon_{i}\\
\beta_{0j} & \sim N(\mu_0, \sigma_0^2)
\end{aligned}
$$

Where $i = 1, 2, \cdots, N$ is the index for individual observations, so the total sample size is $N$ in this case. $j = 1, 2, \cdots, J$ is the index for groups, so there are $J$ groups/clusters in this case. $\beta_{01}, \beta_{02}, \cdots, \beta_{0J}$ are the random intercepts in this model. $\beta_1, \beta_2, \cdots, \beta_K$ are the fixed-effects parameters, and in this case, $K = 3$. The number of predictor variables can vary, and this data generating process can be generalized using a more succinct matrix form style as:
$$
\begin{aligned}
y_{ij} & = \beta_{0j} + \mathbf{X\beta} + \epsilon_{i}\\
\beta_{0j} & \sim N(\mu_0, \sigma_0^2)
\end{aligned}
$$
where $\mathbf{X}$ is a $N\times K$ data matrix, and $\mathbf{\beta}$ is a $K \times 1$ column vector.

We assume that the random intercepts $\beta_{0j}$ has a normal distribution with two hyperparameters: mean $\mu_0$ and standard deviation $\sigma_0$.

Simulation parameters
---



```{r SimulateData, eval=FALSE}
set.seed(666)

Total_obs = 500
group = 10
Group_Obs = Total_obs/group

b1 = 1
mu = 1
sigma = 1
epsilon = 2

# y = b0 + b1 + e
j = rep(1:group, Group_Obs)
b0 = rnorm(group, 1, 5)
x = rnorm(Total_obs, 5, 5)

y = rep(b0, Group_Obs) + b1*x + rnorm(Total_obs)

dat = data.frame(y, j, x, b0, b1)

plot(x, y)
for (i in 1:group) {
  abline(b0[i], b1)
}

reg1 <- lm(y ~ x,data = dat) 
abline(reg1, col = "grey", lwd = 3)

reg2 <- lm(y ~ factor(j) + x,data = dat)
reg2_b0 = reg2$coefficients[labels(reg2$coefficients) == "(Intercept)"]
reg2_b1 = reg2$coefficients[labels(reg2$coefficients) == "x"]
abline(reg2_b0, reg2_b1, col = "blue", lwd = 3)


require(lme4)
reg3 <- lmer(y ~ x + (1|j), dat)
summary(reg3)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
require(ggplot2)

nDriver = 50
Hour = 10

x = seq(0, Hour, 0.1)
driverID = 1:nDriver
b0 = rnorm(nDriver, -2, 1)
b1 = rnorm(nDriver, 0.25, 0.2)

dat = data.frame(driverID = rep(driverID, each = length(x)), 
                 b0 = rep(b0, each = length(x)), 
                 b1 = rep(b1, each = length(x)), 
                 x = rep(x, nDriver))


dat$p = 1 - exp(-exp(dat$b0 + dat$b1*dat$x))

ggplot(data = dat, aes(x = x, y = p, group = driverID)) + 
  geom_line(size = 0.1) + 
  xlab("Cumulative hours of driving") + 
  ylab("Probability of at least 1 critical event") + 
  ggtitle("Random intercept and random slope model of driver fatigue")
```





