---
title: "Data Analysis - The data.table way"
author: "Miao Cai^[Department of Epidemiology and Biostatistics, College for Public Health and Social Justice, Saint Louis University. Email: [miao.cai@slu.edu](miao.cai@slu.edu)]"
date: "1/10/2019"
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
  pdf_document:
    number_sections: yes
link-citations: yes
link-color: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# data.table novice

## Introduction

General form maps to SQL:

- i -> WHERE
- j - > SELECT
- by -> GROUP BY

## Select columns in j

```
mtcarsDT[
  mpg > 20,
  .(AvgHP = mean(hp),
    "MinWT(kg)" = min(wt*453.6)),
  by = .(cyl, under5gears = gear < 5)
]
```

`DT[i, j, by]` is pronounced as: 

> Take `DT`, subset by row using `i`, then calculate `j` grouped by `by`.

`DT[, .(Total = sum(A), Mean = mean(C))]`

## Doing j by group

`DT[, .(MySum = sum(B)), by = .(Grp = A%%2)]`



Group_by summarize and rename the variables:

`DT[, .(Count = .N), by = .(Area = 10 * round(Sepal.Length * Sepal.Width / 10))]`

# data.table yeoman

## Chaining

```{r, eval = FALSE}
ans = DT[, sum(B), by = A]
ans = ans[order(A)]

# by chaining:
ans = DT[, sum(B), by = A][order(A)]
```

## Subset of data - `.SD`

The data.table package provides a special built-in variable `.SD`. It refers to the subset of data for each unique value of the by argument. That is, the number of observations in the output will be equal to the number of unique values in by.

`DT[, lapply(.SD, mean), by = x]`

## subset of data - `.SDcols`

`.SDcols` specifies the columns of DT that are included in .SD. Using .SDcols comes in handy if you have too many columns and you want to perform a particular operation on a subset of the columns (apart from the grouping variable columns).

Using `.SDcols` allows you to apply a function to all rows of a data.table, but only to some of the columns. For example, consider the dog example from the last exercise. If you wanted to compute the average weight and age (the second and third columns) for all dogs, you could assign .SDcols accordingly

```
dogs[, lapply(.SD, mean), .SDcols = 2:3]
   Weight Age
1:     56 5.2
```

Select all but the first row of groups 1 and 2, returning only the grp column and the Q columns

`DT[, .SD[-1], by = grp, .SDcols = paste0("Q", 1:3)]`

```
# Sum of all columns and the number of rows
DT[, c(lapply(.SD, sum), .N), by = x, .SDcols = c("x", "y", "z")]

# Cumulative sum of column x and y while grouping by x and z > 8
DT[, lapply(.SD, cumsum), by = .(by1 = x, by2 = z > 8), .SDcols = c("x", "y")]
```

## add and remove columns

`:=` is defined for use in j only, and is used to update data.tables by reference. One way of using `:=` is the `LHS := RHS` form, where LHS is a character vector of columns (referenced by name or number) you wish to update and RHS is the corresponding value for each column (Note: LHS stands for "left hand side" and RHS stands for "right hand side" in what follows).

```{r}
require(data.table)
t = data.table(mtcars[1:5,])
t

# add columns
t[,c("x1", "x2") := .( am + gear, gear + carb)]; t

t[, `:=` (
  y1 = am - gear,
  y2 = gear + carb
)] ; t


t[,c("mpg", "cyl", "disp") := NULL];t
cols = c("hp", "drat", "x1", "x2")
t[, (cols) := NULL]; t
```

You've had practice with using `:=` in the `LHS := RHS` form. The second way to use `:=` is with functional form:

`DT[, `:=`(colA = colB + colC)]`

Notice that the `:=` is surrounded by two tick marks! Otherwise data.table will throw a syntax error. It is also important to note that in the generic functional form above, `my_fun()` can refer to any function, including the basic arithmetic functions. The nice thing about the functional form is that you can get both the RHS alongside the LHS so that it's easier to read.

## Using set

```
# change the names of "y" to "z"
setnames(DT, "y", "z")

# change the order of columns
setcolorder(DT, c("y", "x"))
```

The `set()` function is used to repeatedly update a data.table by reference. You can think of the `set()` function as a loopable, low overhead version of the `:=` operator, except that `set()` cannot be used for grouping operations. The structure of the `set()` function looks like this:

The function takes four arguments:

- A data.table with the columns you wish to update
- The index used in a loop (e.g. the i in for(i in 1:5))
- The column or columns you wish to update in the loop
- How the column or columns should be updated

For example, Loop through columns 2, 3, and 4, and for each one, select 3 rows at random and set the value of that column to NA

`for (i in 2:4) set(DT, sample(10, 3), i, NA)`

A summary of the set() family:

- `set()` is a loopable, low overhead version of `:=`.
- You can use `setnames()` to set or change column names.
- `setcolorder()` lets you reorder the columns of a data.table.

# data.table expert


```{r}
iris <- data.table::as.data.table(iris)

# Remove the "Sepal." prefix
data.table::setnames(iris, gsub("^Sepal\\.", "", names(iris)))

# Remove the two columns starting with "Petal"
iris[, grep("^Petal", names(iris)) := NULL]
```

## Selecting groups or parts of groups

The previous exercise illustrated how you can manually set a key via `setkey(DT, A, B)`. `setkey()` sorts the data by the columns that you specify and changes the table by reference. Having set a key will allow you to use it, for example, as a super-charged row name when doing selections. Arguments like mult and nomatch then help you to select only particular parts of groups.

Furthermore, two of the instructions will require you to make use of `by = .EACHI`. This allows you to run j for each group in which each item in i joins too. The last instruction will require you to produce a side effect inside the j argument in addition to selecting rows.

```{r}
# The 'keyed' data.table DT
DT <- data.table(A = letters[c(2, 1, 2, 3, 1, 2, 3)], 
                 B = c(5, 4, 1, 9, 8, 8, 6), 
                 C = 6:12)
setkey(DT, A, B)

# Select the "b" group
DT["b"]

# "b" and "c" groups
DT[c("b", "c")]

# The first row of the "b" and "c" groups
DT[c("b", "c"), mult = "first"]

# First and last row of the "b" and "c" groups
DT[c("b", "c"), .SD[c(1, .N)], by = .EACHI]
```



## Rolling joins

To carry the first observation backwards, set the rollends argument to TRUE

```{r}
# Keyed data.table DT
DT <- data.table(A = letters[c(2, 1, 2, 3, 1, 2, 3)], 
                 B = c(5, 4, 1, 9, 8, 8, 6), 
                 C = 6:12, 
                 key = "A,B")

# Print the sequence (-2):10 for the "b" group
DT[.("b", (-2):10)]

# Add code: carry the prevailing values forwards
DT[.("b", (-2):10), roll = TRUE]

# Add code: carry the first observation backwards
DT[.("b", (-2):10), roll = TRUE, rollends = TRUE]
DT[.("b", (-2):10), roll = TRUE, rollends = c(TRUE, TRUE)] # also OK
```

























