---
title: "Data manipulation with data.table"
author: "Miao Cai^[Department of Epidemiology and Biostatistics, College for Public Health and Social Justice, Saint Louis University. Email: [miao.cai@slu.edu](miao.cai@slu.edu)]"
date: "1/10/2019"
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
  pdf_document:
    number_sections: yes
link-citations: yes
link-color: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to data.table

`indices(DT)` makes filtering much faster in data.table

- %like% allows you to search for a pattern in a character or a factor vector
  - Usage: col %like% pattern
- %between% allows you to search for values in the closed interval [val1, val2]
  - Usage: numeric_col %between% c(val1, val2)
- %chin% is similar to %in%, but it is much faster and only for character vectors
  - Usage: character_col %chin% c("val1", "val2", "val3")

These operations are internally parallelized to enable faster speed.

```{r eval = FALSE}
# Subset all rows where start_station starts with San Francisco
batrips[start_station %like% "^San Francisco"]
# Instead of 
batrips[grepl("^San Francisco", start_station)]

# Subset all rows where duration is between 2000 and 3000
batrips[duration %between% c(2000, 3000)]
# Instead of 
batrips[duration >= 2000 & duration <= 3000]

# Subset all rows where start_station is 
# "Japantown", "Mezes Park" or "MLK Library"
batrips[start_station %chin% c("Japantown", "Mezes Park", "MLK Library")]
# much faster than
batrips[start_station %in% c("Japantown", "Mezes Park", "MLK Library")]
```


# Selecting rows from data.table

Selecting a column from a data.frame, you will get a vector, while you will still get a data.table if you select a single column from a data.table using the variable name without quotation.


## Computing on columns

Since columns can be refered to as variables, you can omputing directly on them in `j`.

# Groupwise operations

`.SD` together with `.SDcols` is an incredibly powerful feature that makes computing on multiple columns so much easier.

- `.SD` is a special symbol which stands for Subset of Data
- `.SDcols` holds the columns that should be included in `.SD`

```{r eval = FALSE}
# For each month, find the row corresponding to the shortest trip (by using which.min() on duration).
relevant_cols <- c("start_station", "end_station", 
                   "start_date", "end_date", "duration")

# Find the row corresponding to the shortest trip per month
shortest <- batrips[, .SD[which.min(duration)], 
                    by = month(start_date), 
                    .SDcols = relevant_cols]
shortest

# Find the total number of unique start stations and zip codes per month
unique_station_month <- batrips[, lapply(.SD, uniqueN), 
                                by = month(start_date), 
                                .SDcols = c("start_station", "zip_code")]
unique_station_month
```



# Reference Semantics
## data.table internals

- data.table updates columns in place, i.e., by reference
- This means, you don't need the assign the result back to a variable
- No copy of any column is made while their values are changed
- data.table uses a new operator := to add/update/delete columns by reference

```{r eval=FALSE}
batrips[, c("is_dur_gt_1hour", "week_day") := list(duration > 3600, wday(start_date)]

batrips[, `:=`(is_dur_gt_1hour = NULL,                  
               start_station = toupper(start_station))]
```

As mentioned in the video, data.table does not have to deep copy objects in order to modify them. It can modify objects by reference (inplace) to save memory and runtime using the := operator. When you update a data.table by reference, you don't have to assign the result to a new object as the original data.table is modified directly.

## Adding multiple columns in data.table

More often than not, you may want to add/update multiple columns by reference. There are two ways to do this in data.table:

- LHS := RHS form
- Functional form

In this exercise, you will add columns using the LHS := RHS form. In the LHS, you specify column names as a character vector and in the RHS, you specify values/expressions to be added inside list() (or the alias, .()).


# Read and write data in data.table

Fast data reading with `fread()`

- `select` and `drop` in `fread()` can help choose or drops the columns you want. 
- `skip` can be used to skip meta data.
- `na.strings = c("###", "NA", "9999")`
- `fill = TRUE` to fill in empty cells

```{r eval = FALSE}
data.table::fread("data.csv", nrow = 1000)

# Import using fread
import_fread <- fread("sample.csv", 
                      colClasses = list(factor = 1:4, numeric = 7:10))
str(import_fread)
```

Fast data writing with `fwrite()`

- fwrite() provides three additional ways of writing date and datetime format - ISO, squash and epoch
- Encourages the use of ISO standards with ISO as default


Date and dateimes (squash)

- squash writes yyyy-mm-dd hh:mm:ss as yyyymmddhhmmss, for example.
- Read in as integer. Very useful to extract month, year etc by simply using modulo arithmetic. e.g., 20160912 %/% 10000 = 2016
- Also handles milliseconds (ms) resolution.
- POSIXct type (17 digits with ms resolution) is automatically read in as integer64 by fread.

```{r eval = FALSE}
fwrite(batrips_dates, file = "iso.txt", dateTimeAs = "ISO")
fwrite(batrips_dates, file = "squash.txt", dateTimeAs = "squash")
fwrite(batrips_dates, file = "epoch.txt", dateTimeAs = "epoch")
```


Date and dateimes (Epoch)

- epoch counts the number of days (for dates) or seconds (for time and datetime) since relevant epoch
- Relevant epoch is 1970-01-01, 00:00:00 and 1970-01-01T00:00:00Z for date, time and datetime, respectively
































