---
title: "Visualizing Big Data with Trelliscope"
author: "Miao Cai <miao.cai@slu.edu>"
date: "1/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


# General strategies for visualizing big data
## Binning Continuous Variables using geom_hex()

We would assume that there is a relationship between the total cab fare and the duration of the trip. Since there are too many data points to make a scatterplot, let's use a hexagon-binned plot to investigate this relationship.

```{r}
library(ggplot2)

# Create a hexagon-binned plot of total_amount vs. trip_duration
ggplot(tx, aes(trip_duration, total_amount)) +
  geom_hex(bins = 75) +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
library(dplyr)
library(ggplot2)

# Summarize taxi rides count by payment type, pickup date, pickup day of week
daily_count <- tx %>%
  filter(payment_type %in% c("Card", "Cash")) %>%
  group_by(pickup_date, pickup_dow, payment_type) %>%
  summarise(n_rides = n())

# Plot the data
ggplot(daily_count, aes(pickup_date, n_rides)) +
  geom_point() +
  facet_grid(payment_type ~ pickup_dow) +
  coord_fixed(ratio = 0.4)
```

## Cash / Card Distribution Comparison Using a Quantile Plot

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

ggplot(tx_pop, aes(sample = total_amount, color = payment_type)) +
  geom_qq(distribution = stats::qunif) +
  ylim(c(3, 20))

# Get data ready to plot
amount_compare <- tx_pop %>%
  mutate(total_no_tip = total_amount - tip_amount) %>%
  select(total_amount, total_no_tip, payment_type) %>%
  gather(amount_type, amount, -payment_type)

# Quantile plot
ggplot(amount_compare, aes(sample = amount, color = payment_type)) +
  geom_qq(distribution = stats::qunif, shape = 21) +
  facet_wrap(~ amount_type) +
  ylim(c(3, 20))
```

# trelliscopejs package in R

The trelliscope package is more like a shiny app, constructed by a Javascript API trelliscope.

```{r}
library(ggplot2)
library(gapminder)
library(trelliscopejs)

library(ggplot2)
library(trelliscopejs)

# Create the plot
ggplot(gapminder, aes(year, lifeExp)) +
  geom_line() +
  # Facet on country and continent
  facet_trelliscope(~ country + continent)
```

## Customizing the gapminder display

With the linear model layer added to the plot, specify that automatic cognostics should be computed for this display and explore what new cognostics this layer adds to the display.

```{r}
library(trelliscopejs)
library(ggplot2)

# Create the plot
ggplot(gapminder, aes(year, lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_trelliscope(~ country + continent,
    name = "lifeExp_by_country",
    desc = "Life expectancy vs. year for 142 countries.",
    nrow = 1, ncol = 2,
    # Set the scales
    scales = "sliced",
    # Specify automatic cognistics
    auto_cog = TRUE)
```

All variables in the data passed in to ggplot() are inspected for use as cognostics.

- If the variable is numeric and varies within each panel group, a set of summary statistics is computed for each panel.
- If the variable is constant within each panel group, a single cognostic with that value is computed for each panel.

## Customizing custom cognostics

A function cog() can be wrapped around a variable to fine-tune how a cognostic is handled in Trelliscope.

With cog(), some of the most useful things you can specify include:

- desc: a meaningful description for the cognostic
- default_label: a boolean specifying whether the cognostic should be shown as a label by default or not

```{r}
library(ggplot2)
library(dplyr)
library(gapminder)
library(trelliscopejs)
space_to_dash <- function(x) gsub(" ", "-", x)

# Group by country and create the two new variables
gap <- gapminder %>%
  group_by(country) %>%
  mutate(
    delta_lifeExp = tail(lifeExp, 1) - head(lifeExp, 1),
    ihme_link = paste0("http://www.healthdata.org/", space_to_dash(country)))

# Add the description
gap$delta_lifeExp <- cog(gap$delta_lifeExp, desc = "Overall change in life expectancy")
# Specify the default label
gap$ihme_link <- cog(gap$ihme_link, default_label = TRUE)

ggplot(gap, aes(year, lifeExp)) +
  geom_point() +
  facet_trelliscope(~ country + continent,
    name = "lifeExp_by_country",
    desc = "Life expectancy vs. year.",
    nrow = 1, ncol = 2,
    scales = c("same", "sliced"))
```

```{r eval=FALSE}
library(dplyr)
library(tidyr)
library(purrr)

# Nest stocks by symbol
by_symbol <- stocks %>%
  group_by(symbol) %>%
  nest()

min_volume_fn <- function(x) min(x$volume)

# Create new column
by_symbol_min <- by_symbol %>%
  mutate(min_volume = map_dbl(data, min_volume_fn))
```

an "open-high-low-close" plot using plotly, then create a trelliscope display by wrapping the result with trelliscope() and provide it the name of "ohlc_top500".

```{r}
library(trelliscopejs)
library(dplyr)
library(purrr)
library(plotly)

ohlc_plot <- function(d) {
  plot_ly(d, x = ~date, type = "ohlc",
    open = ~open, close = ~close,
    high = ~high, low = ~low)
}

by_symbol_plot <- by_symbol %>%
  mutate(panel = map_plot(data, ohlc_plot))

trelliscope(by_symbol_plot, name = "ohlc_top500")
```



























































