---
title: "An identification problem in DID"
author: "Miao Cai"
date: "12/20/2018"
output:
  pdf_document:
    number_sections: yes
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

In classic difference-in-difference (DID) settings, there are only two periods, before and after.

$$Y = \beta_0 + \beta_1 \cdot Tr + \beta_2 \cdot Post + \beta_3 T \cdot Post + \epsilon$$

Where $T$ is an indicator of treatment or not, $Post$ is an indicator of pre or post periods, and $\beta_3$ is the DID estimator of treatment causal effect. 


```{r twpperiods}
set.seed(666)

nunits = 400
nyear = 2

b3 = 5
b1 = 3
b2 = 5
b0 = 0

Group = rep(1:nunits, each = nyear)
Year = rep(2007:(2007+nyear-1), times = nunits)
Tr = rep(sample(0:1, nunits, replace = TRUE), each = nyear)
Post = ifelse(Year<2008, 0, 1)



Y = b0 + b1*Tr + b2*Post + b3*Tr*Post  + rnorm(nunits*nyear, 0, 1)

dat = data.frame(Y, Group, Tr, Post)

require(tidyverse)

dat %>%  
  ggplot(aes(x = Post, y = Y, group = Group, color = 
               factor(Tr))) + 
  geom_line(alpha = 0.3)+ 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) + 
  theme_bw()
```



In modern econometrics, long panel data are more and more commonly used. Researchers typically add year dummy variables to allow for time trend fixed effects in their models. 

$$Y = \beta_0 + \beta_1 \cdot Tr + \beta_2 \cdot Post + \beta_3 \cdot T \cdot Post + \beta_4 \cdot Year + \epsilon$$

However, identification may be a potential issue in this case since $Post$ and $Year$ dummies have a lot in common. In this demonstration, I will use simulation to test the identification issue in this scenario.

```{r longpanel}
set.seed(666)

nunits = 400
nyear = 10

b0 = 1
b1 = 2
b2 = 3
b3 = 4

Group = rep(1:nunits, each = nyear)
Year = rep(2007:(2007+nyear-1), times = nunits)
Tr = rep(sample(0:1, nunits, replace = TRUE), each = nyear)
Post = ifelse(Year<2013, 0, 1)

byear = c(0, rnorm(nyear-1, 0, 1))
yyear = model.matrix( ~ YEAR-1, data=data.frame(YEAR = factor(Year))) %*% byear

Y = b0 + b1*Tr + b2*Post + b3*Tr*Post + yyear + rnorm(nunits*nyear, 0, 1)

dat = data.frame(Y, Group, Tr, Post, Year)

require(tidyverse)

dat %>%  
  ggplot(aes(x = Year, y = Y, group = Group, color = factor(Tr))) + 
  geom_line(alpha = 0.3) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) + 
  theme_bw()
```

- $nobs$: `r nunits*nyear`
- $Year$: 2007 - 2016, in total 10 years
- $Tr$: 0 or 1
- $Post$: 0 or 1

- $\beta_0$ = `r b0`
- $\beta_1$ = `r b1`
- $\beta_2$ = `r b2`
- $\beta_3$ = `r b3`

```{r did}
didfit = lm(Y ~ Tr + Post + Tr*Post + factor(Year), data = dat)
summary(didfit)

namerow = rownames(summary(didfit)$coefficients)

estimates = round(summary(didfit)$coefficients[namerow %in% c("(Intercept)", "Tr", "Post", "Tr:Post"), 1], 2)
parameters = c(b0, b1, b2, b3)
```

- The parameters are `r parameters`
- The estimates are `r estimates`

It seems that there are identification issues here: the estimate of $\beta_2$, which is estimates the post-treatment fixed effect, is biased by one unit. The model cannot generate estimate for the last year's fixed effect.
