---
title: "Joining Data in R with data.table"
author: "Miao Cai^[Department of Epidemiology and Biostatistics, College for Public Health and Social Justice, Saint Louis University. Email: [miao.cai@slu.edu](miao.cai@slu.edu)]"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
  pdf_document:
    number_sections: yes
link-citations: yes
link-color: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Joining multiple data.tables

The tables() function will show you all data.tables loaded in your R session

## Joins

Concept of joins come from database query languages (e.g. SQL).

- inner (default): Only keep observations that have information in both data.tables
- full: Keep all observations that are in either data.table
- left: Add information from the right data.table to the left data.table
- right: Add information from the left data.table to the right data.tablea

All these joins can be acheived using the `merge()` from data.table package.

```{r}
require(data.table)
merge(x = demographics, y = shipping, by.x = "name", by.y = "name") # inner join
merge(x = demographics, y = shipping, by.x = "name", by.y = "name", all = TRUE) # full join
merge(x = demographics, y = shipping, by = "name", all.x = TRUE) # left join
merge(x = demographics, y = shipping, by = "name", all.y = TRUE) # right join
```

# Join tables using data.table syntax

The default format for joining in data.table syntax is right join. The general form of data.table syntax joins:

`DT[i, on]`

```{r}
demographics[shipping, on = .(name)]
demographics[shipping, on = list(name)]
demographics[shipping, on = "name"]
```

**Inner-join**: `DT[i, on, nomatch = 0]`

Not possible with the data.table syntax, use the merge() function:

**Anti-join**: Filter a data.table to rows that have no match in another data.table

```{r}
demographics[!shipping, on = .(name)]
```

## Setting data.table keys

- Setting keys means you don't need the on argument when performing a join
  - Useful if you need to use a data.table in many different joins
- Sorts the data.table in memory by the key column(s)
  - Makes filtering and join operations faster
- Multiple keys can be set

```{r}
setkey(DT, key1, key2, key3)
setkey(DT, "key1", "key2", "key3")

keys <- c("key1", "key2", "key3")
setkeyv(dt, keys)
haskey(dt) # haskey() checks whether you have set keys
key(dt) # returns the key columns you have set
```

## Chaining data.table expressions

`DT1[DT2, on][i, j, by]`


# Diagnosing and Fixing Common Join Problems

## Complex keys

> Error in bmerge(i, x, leftcols, rightcols, io, xo, roll, rollends, 
> nomatch,  : 
>  typeof x.age (double) != typeof i.name (character)

The keys are of different data types.

Avoid misspecified joins: Learning what each column represents before joins will help you avoid errors.




































