---
title: |
       BST 5230 Bayesian Statistics - Spring 2018 
       Mid-term Exam
author: "Miao Cai <miao.cai@slu.edu>"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{amsmath}
  - \usepackage[UTF8]{ctex}
  - \usepackage{enumitem}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{afterpage}  
  - \newcommand{\benum}{\begin{enumerate}[label=(\alph*)]}
  - \newcommand{\eenum}{\end{enumerate}}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = 'H')
```

\textbf{1.(10 Points 5,5)) }A disease has base rate 3%.  The sensitivity of a test is 0.80 and the specificity is 0.995.
 \benum
  \item Find the probability that a person selected at random tests positive for the disease.
  \item Find the positive predictive value.
 \eenum

\textbf{Answers:}
\benum
\item \begin{align*}
P(T = +) & = P(T=+,D=+) + P(T=+,D=-)\\
& = P(T=+|D=+)*P(D=+) + P(T=+|D=-)*P(D=-)\\
& = P(T=+|D=+)*P(D=+) + [1-P(T=-|D=-)]*[1-P(D=-)]\\
& = 0.8*0.03+(1-0.995)(1-0.03)\\
& = `r 0.8*0.03+(1-0.995)*(1-0.03)`
\end{align*}
\item \begin{align*}
PPV & = P(D=+|T=+)\\
& = \frac{P(D=+)P(T=+|D=+)}{P(T=+)}\\
& = \frac{0.03*0.80}{0.02885}\\
& = `r 0.03*0.80/0.02885`
\end{align*}
\eenum

\newpage

\textbf{2. (20 Points) }Suppose that $X_1, X_2, \dots X_n \sim \text{i.i.d GEOMETRIC}(\theta)$. That is, they are geometric random variables.  (That is, each is the number of trials needed to get one success where the probability of success is $\theta$.)  The probability mass function for the geometric distribution is $$f(x|\theta) = \theta(1-\theta)^{x-1}, \quad x = 1, 2, 3, \dots $$ Suppose that we assume a BETA($\alpha, \beta$) prior distribution for $\theta$.

 \benum
  \item Find the posterior distribution for $\theta$.
  \item Is the BETA($\alpha, \beta$) prior distribution a conjugate prior?  Explain.
  \item Suppose that $n=5$ and we obtain $x_1=1, x_2 = 5, x_3 = 4, x_4 = 2, x_5 = 8$.  Assume a BETA(2,3) prior distribution.  Find the posterior distribution and plot it.
  \item Find the posterior mean and standard deviation.
 \eenum

\textbf{Answers:}
\benum
\item \begin{align*}
P(\theta|X) & = \frac{c_1\theta^{\alpha - 1}(1-\theta)^{\beta-1}*\theta(1-\theta)^{x-1}}{c_2}\\
& = c_3\theta^{\alpha}(1-\theta)^{x+\beta-1} \quad c_1, c_2, c_3 \text{are supposed to be unknow constants}\\
& \sim \text{BETA}(\alpha + 1, \beta + x - 1)
\end{align*}

\item Yes, BETA($\alpha, \beta$) is a conjugate prior for geometric data because if we have prior BETA($\alpha, \beta$) and geometric data, we will get posterior BETA($\alpha + 1, \; \beta+X-1$), which is in the same family of prior distribution.

\item According to the result from part (a), we have:
\begin{align*}
P(\theta|X) & \sim \text{BETA}(\alpha + 1, \beta + \sum_{i=1}^5x_i - 1)\\
& \sim \text{BETA}(3, 22)
\end{align*}
The plot of the posterior distribution is as follows:
```{r 2.c}
x = seq(0, 1, 0.0001)
d = dbeta(x, 3, 22)
plot(x, d, type = "l", main = "2.(c) Posterior ~ BETA(3, 22)", ylab = "density")
```

\item According to the properties of BETA distribution:
\begin{align*}
Mean & = \frac{\alpha}{\alpha + \beta} = \frac{3}{3 + 22} = \frac{3}{25}\\
s.d. & = \sqrt{Var} = \sqrt{\frac{\alpha \beta}{(\alpha + \beta)^2(\alpha + \beta + 1)}} = \sqrt{\frac{3*22}{(3 + 22)^2(3 + 22 + 1)}} = 0.0637302 
\end{align*}
\eenum

\newpage

\textbf{3. (6 Points) }Choose and plot a prior that describes your belief in the following circumstances.  Explain your reasoning in both cases.

 \benum
  \item Suppose you have a coin that you know is minted by the US government and has not been tampered with.  Therefore you have a strong prior belief that the coin is fair.
  \item Now you have a different coin, this one made of some strange material and marked "Patent Pending: International Magic Company."
 \eenum

\textbf{Answers:}
\benum
\item Since we have a strong belief that the coin should be fair, the mean should be 0.5, the variance should be very small and the support should be within (0, 1). So I choose the prior BETA(10000, 10000). Here is the plot of the prior distribution I choose:
```{r 3.a}
x = seq(0, 1, 0.0001)
d = dbeta(x, 10000,10000)
plot(x, d, type = "l", main = "3.(a) Prior ~ BETA(10000, 10000)", ylab = "density")
```

\item Since we have a strong belief that the coin should be either 0 or 1, the density should be high at either 0 or 1, and the density should be very small in between 0 and 1 and the support should be within (0, 1). So I choose the prior BETA(0.5, 0.5). Here is the plot of the prior distribution I choose:
```{r 3.b}
x = seq(0, 1, 0.0001)
d = dbeta(x, 0.5,0.5)
plot(x, d, type = "l", main = "3.(b) Prior ~ BETA(0.5, 0.5)", ylab = "density")
```
\eenum

\newpage

\textbf{4.  (16 Points) }Suppose that $X|\theta \sim  BIN(4,\theta$), that is binomial with 4 trials and probability of success.  Assume a BETA(2,3) prior distribution for  .  We observed  $x=1$ success out of the 4 trials.

 \benum
  \item Suppose we apply the Metropolis-Hastings algorithm beginning with  $\theta^{(1)}=0.6$.  Suppose we use as the proposal density normal distribution with mean $\theta^{(t)}$  and variance $\sigma^2=0.04$.  Find the acceptance probability if the first proposal is  $\theta^{(\text{prop})} = 0.56$.
  \item Suppose we apply the Metropolis-Hastings algorithm beginning with  $\theta^{(1)}=0.6$.  Suppose we use as the proposal density normal distribution with mean $\theta^{(t)}$  and variance $\sigma^2=0.04$ .  Find the acceptance probability if the first proposal is  $\theta^{(\text{prop})} = 0.62$.  
  \item Suppose we apply the Metropolis-Hastings algorithm beginning with  $\theta^{(1)}=0.6$.  Suppose we use as the proposal density a BETA(2,3) prior distribution.  Find the acceptance probability if the first proposal is $\theta^{(\text{prop})} = 0.56$.
  \item Among (a) – (c) which is (are) the simpler Metropolis and which is (are) the Metropolis-Hastings algorithm.
 \eenum
 
\textbf{Answers:}
\benum
\item The R codes for calculating the acceptance probability and the result are as follows:
```{r 4.a}
# Essential part of BETA distribution
pBeta = function(p,a,b) 
{ 
  z = 0
  if (p >= 0 && p <= 1) {  z = p^(a-1) * (1-p)^(b-1) }
  return(z)
}

# Binomial likelihood function
pLik = function(x,n,p) { p^x * (1-p)^(n-x) }

# Proposal distribution
pProp = function(p)  
{ 
  rnorm(n=1,mean=p,sd=sqrt(0.04)) 
}

n = 4 # the number of trials
x = 1 # the number of success
a = 2 # alpha for beta distribution
b = 3 # beta  for beta distribution

pp = 0.56 # current value
pc = 0.6  # proposal value

num = pBeta(pp,a,b)*pLik(x,n,pp) # numerator
den = pBeta(pc,a,b)*pLik(x,n,pc) # denominator

(alpha = num/den) # acceptance probability - alpha
```

\bigskip

\item The R codes for calculating the acceptance probability and the result are as follows:
```{r 4.b}
pp = 0.62 # current value
pc = 0.6  # proposal value

num = pBeta(pp,a,b)*pLik(x,n,pp) # numerator
den = pBeta(pc,a,b)*pLik(x,n,pc) # denominator

(alpha = num/den) # acceptance probability - alpha
```

\bigskip

\item The R codes for calculating the acceptance probability and the result are as follows:
```{r 4.c}
# Proposal distribution
pProp = function(p)  
{ 
   rbeta(n = 1, shape1 = 2, shape2 = 3) 
}

pp = 0.56 # current value
pc = 0.6  # proposal value

num = pBeta(pp,a,b)*pLik(x,n,pp)*pc*(1-pc)^2 # numerator
den = pBeta(pc,a,b)*pLik(x,n,pc)*pp*(1-pp)^2 # denominator

(alpha = num/den) # acceptance probability - alpha
```

\bigskip

\item In this question, (a) and (b) are the simpler Metropolis algorithm because the proposal density are normal distributions which are symmetric, while (c) is the Metropolis-Hastings algorithm because the proposal density is beta which is not symmetric.


\eenum

