---
title: "Examine the Association between Temperature Variation and Cardiovascular Disease Hospitalization"
subtitle: "A Case Study in Shanxi, China"
author: Miao Cai
institute: |
  | Department of Epidemiology & Biostatistics, 
  | Saint Louis University
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: "boxes"
    colortheme: "whale"
    fonttheme: "structurebold"
    toc: false
    slide_level: 2
csl: numbered.csl
bibliography: refs/ref.bib
header-includes:
  - \usepackage{soulutf8}
  - \usepackage{color}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{lscape}
  - \usepackage{setspace}
  - \usepackage{dcolumn}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Introduction
============
Theoretical background
----------------------
### Global climate change

- Frequent heat shocks and cold spells [@ryti2015global; @huber2017global]
- 12.6 M deaths are attributable to climate change [@watts2017lancet]

### CVD

- Cardiovascular disease (CVD) is the leading cause of death worldwide and in China 
- 15.2 M killed by CVD, which accoutns for 26.7% of all deaths [@who2018]


Research Question
-----------------

Is **\textcolor{red}{temperature variation}** associated with **\textcolor{red}{CVD hospitalizations}**?

- biological association

- research gap: no epidemiological evidence

Data source
===========
Shanxi Province, China
----------------------

```{r fig.cap="The location of Shanxi Province in China"}
knitr::include_graphics("figs/shanxi.pdf")
```

Shanxi inpatient database
-------------------------
::: columns
:::: column

### CVD inpatients
  + 175 hospitals
  + 95% of administrative districts (ADs)
  + ~ 1.2 million CVD patients
  + 2015 - 2017
::::

:::: column
```{r fig.cap="Geographic distribution of 175 hospitals in Shanxi, China"}
knitr::include_graphics("figs/hospitals.pdf")
```
::::
:::

The number of CVD patients by month
-----------------------------------

```{r fig.cap="The number of CVD patients by year and month"}
knitr::include_graphics("figs/Npat_by_month.pdf")
```




## Temperature variation definition

- Vertical variation: $V_v = T_{max} - T_{min}$
- Horizontal variation: $V_h = \sum_{i = \text{d} - 7}^{\text{d}} \Big|T_i - T_{i - 1}\Big|$

```{r fig.cap="Maximum and minimum temperature in 118 ADs in Shanxi, 2015 - 2017"}
knitr::include_graphics("figs/temperature_in_groups.pdf")
```

## Temperature variation trend

```{r fig.cap="Temperature variation trend in 118 ADs in Shanxi, 2015 - 2017", fig.width=5}
knitr::include_graphics("figs/two_tem_variation.pdf")
```

Shanxi Statistical Yearbook
---------------------------

::: columns
:::: column
### County/city level
  + 118 ADs
  
### Variables available
  + Population
  + Gender
  + Rural
  + GDP
  + Others
::::

:::: column
```{r fig.cap="Choropleth map patient distribution in each county and city in Shanxi, China"}
knitr::include_graphics("figs/choropleth_patients.pdf")
```
::::
:::

## Final sample

- 109 ADs
- 12 months in 3 years
- $109*12*3 = 3,924$ rows

Data cleaning, visualization, statistical models, and reporting are conducted in \textcolor{blue}{\texttt{R 3.6.0}}.

# Statistical models
## Model selection

The outcome is the count of CVD patients: 

- Must be \textcolor{red}{non-negative}, 
- \textcolor{red}{integers}, 
- \textcolor{red}{highly righ-skewed}.

OLS assumpts are **\textcolor{red}{not met}**!

\textcolor{blue}{Poisson regression} is a natural model for count data. We consider:

1. non-spatial Poisson regression
2. random effects Poisson regression
3. spatial Poisson regressions


## Non-spatial Poisson model

$$
\begin{aligned}
Y_i & \sim \text{Poisson}(\text{Pop}_i*\lambda_i)\\
\log(\lambda_i) & = \beta_0 + \sum_{k}\mathbf{\beta X}
\end{aligned}
$$

- \textcolor{blue}{$Y_i$}: The number of CVD inpatients
- \textcolor{blue}{Pop$_i$}: The total population as an \textcolor{red}{offset}
- \textcolor{blue}{$\lambda_i$}: The rate parameter of Poisson distribution, both the mean and vaiance
- Predictor variables: \textcolor{blue}{temperature variation (horizontal and vertical), average temperature, log of GDP per capita, the percent of female, and the percent of rural population}

## Non-spatial Hierarchical Poisson model

Assuming that each ADs have its own characteristics, and observations are **\textcolor{blue}{conditionally independent}**.

$$
\begin{aligned}
Y_{ij} & \sim \text{Poisson}(\text{Pop}_{ij}*\lambda_{ij})\\
\log(\lambda_{ij}) & = \beta_{0j} + \beta_{1j}HV_{ij} + \beta_{2j}VV_{ij} + \sum_{k-2}\mathbf{\beta X} + u_{0j}\\
\beta_{0j} & \sim N(\mu_0, \sigma_0^2)\\
\beta_{01} & \sim N(\mu_1, \sigma_1^2)\\
\beta_{02} & \sim N(\mu_2, \sigma_2^2)\\
\end{aligned}
$$

- \textcolor{blue}{Random intercepts}: $\beta_{0j}$ for each AD $j$

- \textcolor{blue}{Random slopes}: $\beta_{1j}$ and $\beta_{2j}$ for each AD $j$


## Spatial Poisson models

### Spatial lag Poisson model

- An autoregressive parameter $\rho$
- A weighting matrix $\mathbf{W}$

$$
\begin{aligned}
Y_i & \sim \text{Poisson}(\text{Pop}_i*\lambda_i)\\
\log(\lambda_i) & = \beta_0 + \sum_{k}\mathbf{\beta X} + \rho\mathbf{W}y
\end{aligned}
$$


### Geographically weighted Poisson model


Still at the initial proof of concept stage. It can be conducted in the `R` package `spgwr` by Roger Bivand (2017).

Results
=========

## Moran's I - CVD patients

\scriptsize
```{r}
pacman::p_load(dplyr, spdep)
load("data/hloca.Rdata")
names(hloca)[names(hloca) == "longtitude"] = "longitude"
hloca = hloca[hloca$N > 1000,]
```
```{r echo=TRUE}
coords = hloca %>% 
  select(longitude, latitude) %>% 
  as.matrix()
nb = dnearneigh(coords,d1=0, d2 = 70, longlat = TRUE) 
moran.test(hloca$N, nb2listw(nb, style="W"))
```

\normalsize


## non-spatial Poisson models


```{r results='asis', echo = FALSE}
pacman::p_load(stargazer)

load("data/fit1.Rdata")
load("data/fit2.Rdata")

stargazer(fit1, fit2, header = FALSE, font.size='scriptsize', 
          title = "Parameter estimates of non-spatial Poisson models",  
          single.row = TRUE, dep.var.labels=c("Number of CVD hospitalizations"),
          colnames= FALSE,object.names=FALSE,model.numbers = TRUE,
          covariate.labels=c("Between variance", "Within variance", "Mean temperature", 
                             "Log GDP", "Rural", "Female", "Constant"),
          model.names = FALSE,align=TRUE,column.sep.width="-30pt",
          column.labels=c("Poisson","Mixed-effects Poisson"), 
          omit.stat = c("n", "bic"),  dep.var.labels.include = FALSE)
```

## Spatial Poisson models

To be added

- Spatial lag Poisson model
- Geographically weighted Poisson model

Discussion
==========

## Limitations

- Assumptions: the patients did not seek care in other ADs

- Omitted variable bias: air pollution, older population

- Different weights for different severity of CVD hospitalization?


## Future directions


- Add time series analysis in several major cities?

- Different effects in cold and hot days?

- Hierarchical spatial Poisson regression?

## References






