---
title: "Simulating multilevel models in the Matrix form in `stan`"
author: "Miao Cai <miao.cai@slu.edu>"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
    toc: yes
header-includes:
  - \usepackage{float}
  - \usepackage{setspace}
geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.5cm"
linkcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


# Model setting

A simple random intercept model. Let us assume that there are three predictor variables $x_1, x_2, x_3$, and **10 groups** ($d(i) = 1, 2, \cdots, 10$):

$$
\begin{aligned}
y_i & = \beta_{0,d(i)} + \beta_1x_1 + \beta_2x_2 + \beta_3x_3 + \epsilon_i\\
\epsilon_i & \sim N(0, \sigma^2)\\
\beta_{0,d(i)} & \sim N(\mu_0, \sigma_0^2)
\end{aligned}
$$

The goal here is to write this $\beta_1x_1 + \beta_2x_2 + \beta_3x_3$ term in a matrix multiplication form in `Stan`. Here we assume the predictor variables were generated from the following distribution:

$$
\begin{aligned}
x_1 & \sim N(3, 10)\\
x_2 & \sim \text{Gamma}(10, 10)\\
x_3 & \sim \text{Poisson}(10)
\end{aligned}
$$

The parameter settings are:

$$
\begin{aligned}
\text{Hyperparameters:} &\quad \mu_0 = 2, \sigma_0 = 5\\
\text{Fixed parameters:} & \quad \beta_1 = 2.5, \beta_2 = 0.5, \beta_3 = -1.3\\
\text{Variance parameter:} & \quad \sigma = 2
\end{aligned}
$$

\clearpage
# Data simulation
```{r}
set.seed(123)
# Simulation setting
d = 10 # total number of groups
D = rpois(d, 100) # sample size in each group
n = sum(D) # total sample size

# Hyper-parameters
mu0 = 2; sigma0 = 5

# Parameters
B = c(2.5, 0.5, -1.3); sigma = 2

# Random intercepts
b0 = rnorm(d, mu0, sigma0)
B0 = rep(b0, D)

# Predictor variables
id = rep(1:d, D)
x1 = rnorm(n, 3, 10)
x2 = rgamma(n, 10, 10)
x3 = rpois(n, 10)
epsilon = rnorm(n, 0, sigma)
X = matrix(c(x1, x2, x3), nrow = n)

Y = B0 + X %*% B + epsilon
df = data.frame(cbind(id, Y, X))
names(df) = c("id", "y", paste0("x", 1:3))
```

The random intercepts are:

```{r}
df_b0 = data.frame(t(b0))
names(df_b0) = paste0("b0_", 1:d)
knitr::kable(df_b0, digits = 3)
```

\clearpage
# Estimation with `lme`

```{r}
pacman::p_load(lme4, rstanarm, brms, rstan, broom)

fit0 <- lme4::lmer(y ~ x1 + x2 + x3 + (1 | id), data = df)

broom::tidy(fit0) %>% 
  knitr::kable(digits = 3)
```

\clearpage
# Esitmation with `rstanarm`

```{r}
f_rstanarm = stan_lmer(y ~ x1 + x2 + x3 + (1 | id),
                       data = df, 
                       chains = 1, iter = 1000, refresh = 0)
```

## Parameter estimates

```{r}
broom::tidy(f_rstanarm, parameters = "non-varying") %>% 
  knitr::kable(digits = 3)
broom::tidy(f_rstanarm, parameters = "varying") %>% 
  knitr::kable(digits = 3)
broom::tidy(f_rstanarm, parameters = "hierarchical") %>% 
  knitr::kable(digits = 3)
broom::tidy(f_rstanarm, parameters = "auxiliary") %>% 
  knitr::kable(digits = 3)
```

## Default priors

```{r}
prior_summary(f_rstanarm)
```


\clearpage
# Esitmation with `brms`

```{r}
f_brms = brm(y ~ x1 + x2 + x3 + (1 | id), 
             data = df, family = gaussian(), 
             chains = 1, iter = 1000, refresh = 0)
```

## Parameter estimates

```{r}
broom::tidy(f_brms) %>% 
  knitr::kable(digits = 3)
```

## Default priors

```{r}
prior_summary(f_brms)
```

\clearpage
\linespread{0.9}
## Return `stan` code using `brms::stancode()`
```{r comment=NA}
stancode(f_brms)
```


\clearpage
## Return `stan` code using `brms::make_stancode()`
```{r comment=NA}
make_stancode(y ~ x1 + x2 + x3 + (1 | id), data = df, family = gaussian())
```
\linespread{1}

\clearpage
# Self-define R code

```{r}

```

