---
title: "repex - separating date time"
author: "Miao Cai <miao.cai@slu.edu>"
date: "7/4/2019"
output: 
  pdf_document:
    number_sections: true
    toc: yes
header-includes:
  - \usepackage{float}
geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```

# Example data {-}

```{r}
pacman::p_load(data.table, lubridate)

# example data
ex = function(){
  driver = rep(c("foo", "bar"), each = 11L)
  ping_time = ymd_hm(c(
    "2015-05-27 07:11", "2015-05-27 07:25", "2015-05-27 07:35", "2015-05-27 07:42", "2015-05-27 07:53",
    "2015-05-27 08:09", "2015-05-27 08:23", "2015-05-27 08:39", "2015-05-27 08:52", "2015-05-27 09:12", "2015-05-27 10:25",
    "2015-05-27 16:12", "2015-05-27 16:31", "2015-05-27 16:39", "2015-05-27 16:53", "2015-05-27 17:29",
    "2015-05-27 17:41", "2015-05-27 17:58", "2015-05-27 18:09", "2015-05-27 18:23", "2015-05-27 18:43", "2015-05-27 20:51")
  )
  df = data.table(driver, ping_time)
  setkey(df, driver, ping_time)
  df[,ping_id := .I]
  return(df)
}
df = ex()
stargazer::stargazer(df, summary = FALSE, header = FALSE, 
                     title = "Example data", table.placement = "H")
```


Solution 1: floor cumulative time
=================================

## Separate ping using `flag0`

```{r}
ping_df = ex()
ping_df[,diff := as.integer(difftime(ping_time, shift(ping_time, 1),
                                     units = "mins")),
        by = driver]
ping_df[, diff := {diff[1] = 0L; diff}, driver]
ping_df[,cum_mins := cumsum(diff), driver]
ping_df[,cum_halfhour := round(cum_mins/30, 3), driver]
ping_df[,flag0 := floor(round(cum_mins/30, 3)), driver]
setkey(ping_df, driver, ping_time)
ping_df = ping_df[,.(ping_id, driver, ping_time, diff, cum_mins, flag0)]
stargazer::stargazer(ping_df, summary = FALSE, header = FALSE, 
                     title = "Flagged ping data (solution 1)", table.placement = "H")
```

## Construct trips data

```{r}
# construct trip_df
t1 = ping_df[, .SD[c(1)], by = c("driver")]
t2 = ping_df[, .SD[c(.N)], by = c("driver", "flag0")]
trip_df = rbindlist(list(t1, t2), TRUE)
setkey(trip_df, driver, ping_time)
trip_df[,end_time := shift(ping_time, type = "lead", fill = NA), driver]
trip_df = trip_df[!is.na(end_time), ]
trip_df[, trip_id := .I]
trip_df[, trip_time := as.integer(difftime(end_time, ping_time, 
                                          units = "mins"))]
trip_df = trip_df[, .(trip_id, driver, start_time = ping_time, end_time, trip_time, cum_mins)]

stargazer::stargazer(trip_df, summary = FALSE, header = FALSE, 
                     title = "Constructed trips data (solution 1)", table.placement = "H")
```

## Merge trip_df back to ping_df to have a common id

```{r}
# merge trip_id back to ping data
setkey(trip_df, start_time, end_time)
ping_df[,dummy := ping_time]
pitr_df = foverlaps(ping_df, trip_df,
                    by.x = c("ping_time", "dummy"))[, dummy := NULL][
                      ,.(ping_id, driver, ping_time, trip_id, start_time, end_time, trip_time)]

stargazer::stargazer(pitr_df, summary = FALSE, header = FALSE, 
                     title = "Merged ping and trips data (solution 1)", table.placement = "H")
```


Solution 2: round cumulative time to nearest half hour
=====================================================

## Separate ping using `flag0`

```{r}
ping_df = ex()
# 2a. the same as soltion 1
ping_df[,diff := as.integer(difftime(ping_time, shift(ping_time, 1),
                                     units = "mins")),
        by = driver]
ping_df[, diff := {diff[1] = 0L; diff}, driver]
ping_df[,cum_mins := cumsum(diff), driver]
ping_df[,cum_halfhour := round(cum_mins/30, 3), driver]
ping_df[,flag0 := floor(round(cum_mins/30, 3)), driver]
# 2b. further optimize the range
ping_df[,ind := flag0 - shift(flag0, fill = 0), driver]
ping_df[, dif := cum_halfhour - floor(cum_halfhour) -
          (ceiling(shift(cum_halfhour, fill = 0)) - shift(cum_halfhour, fill = 0)), driver]
ping_df[ind == 0, dif := NA]
ping_df[dif < 0, flag0 := flag0 - 1]
setkey(ping_df, driver, ping_time)
ping_df = ping_df[,.(ping_id, driver, ping_time, diff, cum_mins, flag0)]

stargazer::stargazer(ping_df, summary = FALSE, header = FALSE, 
                     title = "Flagged ping data (solution 2)", table.placement = "H")
```

## construct trips data

```{r}
t1 = ping_df[, .SD[c(1)], by = c("driver")]
t2 = ping_df[, .SD[c(.N)], by = c("driver", "flag0")]
trip_df = rbindlist(list(t1, t2), TRUE)
setkey(trip_df, driver, ping_time)
trip_df[,end_time := shift(ping_time, type = "lead", fill = NA), driver]
trip_df = trip_df[!is.na(end_time), ]
trip_df[, trip_id := .I]
trip_df[, trip_time := as.integer(difftime(end_time, ping_time, 
                                          units = "mins"))]
trip_df = trip_df[, .(trip_id, driver, start_time = ping_time, end_time, trip_time, cum_mins)]

stargazer::stargazer(trip_df, summary = FALSE, header = FALSE, 
                     title = "Constructed trips data (solution 2)", table.placement = "H")
```

## Merge trip_df back to ping_df to have a common id

```{r}
setkey(trip_df, start_time, end_time)
ping_df[,dummy := ping_time]
pitr_df = foverlaps(ping_df, trip_df,
                    by.x = c("ping_time", "dummy"))[, dummy := NULL][
                      ,.(ping_id, driver, ping_time, trip_id, start_time, end_time, trip_time)]

stargazer::stargazer(pitr_df, summary = FALSE, header = FALSE, 
                     title = "Merged ping and trips data (solution 2)", table.placement = "H")
```


Solution 3: Fixed half-hour intervals
=====================================

## Construct trips data

```{r}
ping_df = ex()
trip_df = ping_df[, .SD[c(1, .N)], by = c("driver")]
trip_df[,time_type := rep(c("start_time", "end_time0"), .N/2)]
trip_df[,time_type := factor(time_type, levels = 
                                         c("start_time", "end_time0"))]
trip_df = dcast(trip_df, driver ~ time_type, value.var = "ping_time")
trip_df[,trip_time := as.integer(difftime(end_time0, start_time, 
                                     units = "mins"))]
trip_df[,trip_units := ceiling(trip_time/30)]
trip_df = trip_df[rep(seq(.N), trip_units), !c("trip_time", "trip_units")]
trip_df[,add1 := 0:(.N-1), by = driver]
trip_df[,start_time := start_time[1] + add1*30*60, driver]
trip_df[,end_time := shift(start_time, type = "lead"), by = driver]
trip_df[,end_time := {end_time[.N] = end_time0[.N]; end_time}, 
        by = driver][,c("end_time0", "add1") := NULL]
trip_df[, trip_time := as.integer(difftime(end_time, start_time, 
                                          units = "mins"))]
trip_df[, new_trip_id := .I]
trip_df = trip_df[, .(new_trip_id, driver, start_time, 
                      end_time, trip_time)]
```

```{r}
stargazer::stargazer(trip_df, summary = FALSE, header = FALSE, 
                     title = "Constructed trips data (solution 3)", table.placement = "H")
```

## Merge trip_df back to ping_df to have a common id

```{r}
setkey(trip_df, start_time, end_time)
ping_df[,dummy := ping_time]
pitr_df = foverlaps(ping_df, trip_df,
                 by.x = c("ping_time", "dummy"))[, dummy := NULL][
                   ,.(ping_id, driver, ping_time, new_trip_id, start_time, end_time, trip_time)]
```

```{r}
stargazer::stargazer(pitr_df, summary = FALSE, header = FALSE, 
                     title = "Merged ping and trips data (solution 3)", table.placement = "H")
```

