---
title: "Flowcharts using DiagrammeR"
author: "Miao Cai"
date: "12/20/2018"
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
  pdf_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


Creating flowcharts or conceptual frameworks is a commonly used tool in empirical research. We can create beautifully designed vector flowcharts using [DiagrammeR](http://rich-iannone.github.io/DiagrammeR/io.html) and grViz.

```{r diagramme}
require(DiagrammeR)
require(DiagrammeRsvg)
require(magrittr)

grViz("
      
digraph boxes_and_circles {

  # add note statement
  node [shape = circle
        fontname = Arial
        penwidth = 2.0]
  A; B; C; D; E; F
  
  node [shape = box]
  1; 2; 3; 4; 5; 6; 7; 8
  
  # add edge statements
  # edge [arrowhead = diamond]
  A -> 1; B -> 2; B -> 3; B -> 4; 
  C -> A; 1 -> D; E -> A; 2 -> 4;
  1 -> 5; 1 -> F; E -> 6; 4 -> 6; 
  5 -> 7; 6 -> 7; 3 -> 8[label = 'a label!'
                          fontname = Times]
  
  # add a graph statement
  graph [nodesep = 0.5]
  
}
      
") 
```

The initial statement `digraph boxes_and_circles` is for a directed graph in the DOT language. comment lines begin with a # sign. These lines are disregarded.

`node [shape = circle]` states that all the following nodes to be defined will have the shape of a circle, it could be changed into `box`. `A; B; C; D; E; F` are defining the IDs of each node and calling the `grViz` function repeatedly, and multiple statements can be placed on a single line, separated by a ';'.


Then you can add the directions and relationships of these nodes by using `->` and each separated by a `;`. In order to change a range of properties for the graph, you can add a graph statement. `graph [nodesep = 0.1]` property modifies the separation distance between nodes. 

- change the fonts by adding `fontname = Arial` in square brackets.
- change the stroke width of the nodes by `penwidth = 2.0` in square brackets
- change the appearance of arrows by `arrowhead = diamond` in square brackets
- add the label for edges


```{r mermaid}
mermaid("
        graph LR
        A --> B
        ", height = 200)
```

